Dike experiment

### 2022-12-15 manuscript.Rmd
Update

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  bookdown::word_document2:
    reference_docx: template_word.docx
    global_numbering: TRUE
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

@@

Open Research: Data and code are permanently available on Zenodo under: <https://doi.org/10.XXXX>.

@@

See Figure \@ref(fig:figure-map)

@@

```{r set-up, include = FALSE}
library(bookdown)
library(knitr)
knitr::opts_chunk$set(
  echo = FALSE,
  dpi = 50
  )

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(figure-1)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [@lfu2021waterlevel,@lfu2021discharge] (Table \@ref(table-s1)). Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <mark>Is that right?</mark> and a with an excavator the substrates were mixed <mark>Is that right?</mark> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland (Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016 ). The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <mark>check numbers</mark>, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(table-2))<mark>check</mark>. The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(table-3)) <mark>more details and check</mark>. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(figure-1)). The split plot was cuased by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(table-1)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ). The vegetation was surveyed in June or July 2018--2021 (**Braun-Blanquet, 1928/1964**). Establishment rates of species

@@

We performed all analyses in R (Version 4.2.2; @base), with the functions 'blme' (based on 'lme4') for BLMM [@lme4] **Chung, Rabe-Hesketh, Dorie, Gelman, & Liu, 2013)**; and 'DHARMa' for model evaluation [@DHARMa].

@@

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). <mark>Mention Naturschutzbehörde and funding from WIGES.</mark> We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

@@

Data and code is available on Zenodo: <mark><https://doi.org/10.XXXX></mark>

@@

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020. <mark>funding number WIGES</mark>

@@

@@

## Table 1 {.unnumbered}

```{r table-1, fig.cap = "table-1"}
#include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

```{r table-2, fig.cap = ""}
#include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

@@

```{r figure-1, fig.cap = "\\label{fig:figure-1}This is a caption"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

@@

```{r figure-2, fig.cap = ""}
#knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

@@

```{r figure-3, fig.cap = ""}
#knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-4, fig.cap = ""}
#knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-5, fig.cap = ""}
#knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-6, fig.cap = ""}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

Three dry years (2018--2020) and three minor floods (2018, 2019, 2021) occurred during the study period. Annual temperature, precipitation and discharge of River Danube near the study sites (2017--2021), based on weather station Metten (mean, 1981--2010; ID, 3271; WGS84 (lat/lon), 48.85476 and 12.918911; CDC, 2022a, 2022b) and stream gauge Pfelling (ID, 10078000; WGS84 (lat/lon), 48.87975 and 12.74716; LfU, 2021a). HQ2 = Highest discharge with a probability of occurrence every second year. HSW = Highest water level for shipping.

```{r table-s1}
#knitr::include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

```{r table-s2}
#knitr::include_graphics(here::here("outputs", "tables", "figureS2.png"))
```

@@

```{r table-s3}
#knitr::include_graphics(here::here("outputs", "tables", "table_a3_seedmixes.csv"))
```

>>>>

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

# Start

@@

Open Research: Data and code are permanently available on Zenodo under: <!--# https://doi.org/10.XXXX -->.

@@

@@

```{r set-up, include = FALSE}
library(knitr)
library(bookdown)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 50
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 10
  )

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-1)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [@lfu2021waterlevel,@lfu2021discharge] (Table \@ref(fig:table-s1)). Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland (Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016 ). The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <!--# check numbers  -->, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(fig:table-2))<!--# check  -->. The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(fig:table-3)) <!--# More details and check  -->. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-1)). The split plot was cuased by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(fig:table-1)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ). The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964]. Establishment rates of species

@@

We performed all analyses in R [Version 4.2.2; @base], with the functions 'blme' (based on 'lme4') for BLMM [@lme4, @chung2013]; and 'DHARMa' for model evaluation [@DHARMa].

@@

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) and Klaus Rachl and Stefan Radlmair (Regierung von Niederbayern) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

@@

Data and code is available on Zenodo: <!--# https://doi.org/10.XXXX -->

@@

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020. <!--# funding number of WIGES -->

@@

<!---BLOCK_LANDSCAPE_START--->

@@

\## Table 1

Mischprobe aus mehreren Plots. Measured in \< 2mm fraction <!--# Write full description -->

```{r table-1, fig.cap = "table_1"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  fit_to_width(16.5, inc = 1L, max_iter = 20, unit = "cm")
```

@@

Each plot received a individual set of 20 species with some restrictions to the number of species per functional group. <!--# Full description -->

```{r table-2, fig.cap = "table_2"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(top = FALSE, values = c("", "", "#", "#", "%", "%")) %>%
  autofit()
```

@@

<!---BLOCK_LANDSCAPE_STOP--->

@@

```{r figure-1, fig.cap = "figure-1"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

@@

```{r figure-2, fig.cap = "figure-2"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

@@

```{r figure-3, fig.cap = "figure-3"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-4, fig.cap = "figure-4"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-5, fig.cap = "figure-5"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-6, fig.cap = "figure-6"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

Three dry years (2018--2020) and three minor floods (2018, 2019, 2021) occurred during the study period. Annual temperature, precipitation and discharge of River Danube near the study sites (2017--2021), based on weather station Metten [mean, 1981--2010; ID, 3271; WGS84, lat/lon, 48.85476, 12.918911; @cdc2022prec, @cdc2022temp] and stream gauge Pfelling [ID, 10078000; WGS84, lat/lon, 48.87975, 12.74716; @lfu2021discharge]. HQ2 = Highest discharge with a probability of occurrence every second year. HSW = Highest water level for shipping.

```{r table-s1, fig.cap = "table-s1"}
read_csv(here::here("data", "raw", "data_raw_table_s1.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  compose(part = "body", i = ~ Unit == "m3 s−1", j = "Unit", value = as_paragraph("m", as_sup("3"), " s", as_sup("−1"))) %>%
  autofit()
```

@@

Establishment rate <!--# Full description an include it here -->

```{r table-s2, fig.cap = "table-s2"}
knitr::include_graphics(here::here("outputs", "tables", "table_a2_establishment.html"))
```

@@

See repository outputs/tables/table_a3_seedmixes.csv <!--# Write full link and full description -->
